\newpage
\TOCadd{Acknowledgements}

\begin{center}
ACKNOWLEDGEMENTS
\end{center}

E mihi ana ki ng\=a tohu o nehe, o Te Whanganui-a-Tara e noho nei au.
He mihi ki te iwi e tau nei, Taranaki Wh\=anui me Ng\=ati Toa t\=en\=a koutou.

This research was part of Antarctic Science Platform Project 1: NZARI Ross Ice Shelf Project and was funded by the Royal Society of New Zealand’s Rutherford Discovery Fellowship (Contract: RDF-VUW1602). 
funding support from the New Zealand Ministry of Business and Innovation and Employment through the Antarctic Science Platform contract (ANTA1801) Antarctic Ice Dynamics Project (ASP-021-01).
The equipment used to collect the low-frequency radar profiles was lent by Ginny Catania (Jackson School of Geosciences, University of Texas in Austin). Greg Leonnard and Kelly Gragg (School of Surveying, University of Otago) were instrumental in getting this radar set-up functioning for this work.
Our thanks to Antarctica New Zealand for the field support which was critical to the success of this research. Additionally, VUW Doctoral and Submission Scholarships.

Thank you to all collaborators who supported and enabled this work, Huw Horgan, Wei Ji Leong, Martin Forbes, Alena Malyarenko. 
% Additional thanks to Craig Stevens, Craig Stewart, Karen Alley, Irena Vankova, Dustin Carroll, Oliver Marsh for responding to emails with useful information.
A humongous thanks to both my primary supervisor, Huw Horgan for genuine support, mentorship, encouragement and teaching, and secondary supervisor Brian Anderson, for constructive comments on new topics.

Between waves of lock downs, thanks to the whole crew at the Antarctic Research Centre for fostering such a positive work environment. Thanks to Rob, Dao, Michelle and Aleksander for streamlining running of the ARC, administration and computer systems. Thanks to Meseret and Bereket for keeping the ARC in order, and for conversations and support. Thanks to the other PhD Students for camaraderie and fun times, the originals---Lauren, Wei Ji, and Clarrie, and later---Matt, Fran, Alana. 

Very thankful for ongoing, limitless support from Mum and Dad, Andrew, Emma, Lily and Lola.  It's no stroke of luck that I ended doing a PhD in Antarctic geophysics, with inspiring and supportive parents who've explored the field. Acknowledging the lockdown workmates---Blake, Henry, Bron, Joane and earlier flatties/friends, Laura, Dhaxna, Meg and Isaac, all cordial cooks, and palatable people.   
Thanks to the Wellington hills and beautiful southern coast for perfect runs and swims which fostered a good life. Cheers to the adventure mates, the BSR lads, the WoRM folk, and the TTC descendants, running, rogaining, skiing and biking.

Joane, thanks for the nail-on-the-head advice and engagement on completely new topics, but also---hunkering in the storm-battered tent at Tenneson, overlapping on the Tararua Mountain Race, finding a wh\=eke at Taputeranga, a freezing dip in the Atiwhakatu, or cooking on the fire at M\=akara. 